---
title: "AERA journals WOS general descriptives"
output: 
  html_notebook: 
    toc: yes
    toc_float: true
---

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
library(readr)
library(dplyr)
library(here)
library(tidyr)
library(ggplot2)
library(stringr)
library(directlabels)
library(ggthemes)
library(assertr)
# library(hrbrthemes)


library(knitr)
opts_chunk$set(echo=FALSE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
aera_full = read_csv(here::here('data/processed', '20180419_wos_aera_articledata.csv')) %>%
  verify(dim(.) == c(5919, 73))
#spec(aera)
```

# Published item types

RF 20/04: Ik denk dat we ons best richten op de volgende 4 document types: Article / Article; Book Chapter / Article; Proceedings Paper / Review. De andere types lijken me niet nodig voor de analyses => Filter 800 items.

```{r, message=FALSE, warning=FALSE}
aera_full %>%
  group_by(document_type) %>%
  tally() %>%
  arrange(desc(n)) %>%
  add_tally(n)
```

```{r}
aera = aera_full %>%
  verify(nrow(.) == 5919) %>%
  filter(document_type %in% c('Article', 'Review', 'Article; Book Chapter', 'Article; Proceedings Paper')) %>%
  verify(nrow(.) == 5119)

aera %>% 
  group_by(document_type) %>%
  tally() %>%
  add_tally(n)
```

```{r}
aera %>%
  group_by(document_type, source_title_shortcode) %>%
  tally() %>%
  spread(source_title_shortcode, n)
```

```{r}
aera %>%
  group_by(document_type, publication_year) %>%
  tally() %>%
  spread(document_type, n)
```

# Published item count

## Item counts per AERA-journal

```{r}
aera %>%
  group_by(source_title_shortcode) %>%
  tally() %>%
  rename(total_published_items = n)
```

## Item counts per year 

```{r}
aera %>%
  group_by(publication_year) %>%
  tally() %>%
  rename(total_published_items = n)
```

## Item counts per year and AERA-journal

```{r}
aera %>%
  group_by(publication_year, source_title_shortcode) %>%
  tally() %>%
  spread(source_title_shortcode, n)
```

Year-selection: end time series on 2017, i.e. filter records from 2018 (28 records), as incomplete year gives bias.

```{r}
aera = aera %>%
  verify(nrow(.) == 5119) %>%
  filter(publication_year != 2018) %>% # todo change to character on read-in
  verify(nrow(.) == 5091)
```


```{r, warning=FALSE}
d.aera.journal.year = aera %>%
  group_by(source_title_shortcode, publication_date_full) %>%
  tally()

p.aera.journal.year = ggplot(d.aera.journal.year, aes(
  x = publication_date_full, y = n,
  label = source_title_shortcode,
  group = source_title_shortcode, color = source_title_shortcode))


p.aera.journal.year = p.aera.journal.year + geom_point() + 
  geom_smooth(method = 'loess') + geom_dl(method = 'smart.grid')

p.aera.journal.year = p.aera.journal.year + theme_tufte()  + theme(legend.position = 'none')

p.aera.journal.year + labs(
  title = 'Published item-count is rising, but this is a return to previous levels',
  subtitle = 'Number of items published in AERA-journals per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL
)

# aera %>%
#   group_by(publication_date) %>%
#   tally() %>%
#   mutate()
```

# Published item page count

```{r}
d.aera.journal.year.pgs = aera %>%
  group_by(source_title_shortcode, publication_date_full) %>%
  summarise(
    page_count_n = sum(page_count),
    page_count_mean = mean(page_count),
    page_count_median = median(page_count))

p.aera.journal.year.pgs = ggplot(d.aera.journal.year.pgs, aes(
  x = publication_date_full, y = page_count_n,
  label = source_title_shortcode,
  group = source_title_shortcode, color = source_title_shortcode))


p.aera.journal.year.pgs = p.aera.journal.year.pgs + geom_point() + 
  geom_smooth(method = 'loess') + geom_dl(method = 'smart.grid')

p.aera.journal.year.pgs = p.aera.journal.year.pgs + theme_tufte()  + theme(legend.position = 'none')

p.aera.journal.year.pgs + labs(
  title = 'Total number of pages published is generally risng',
  subtitle = 'Number of pages published in AERA-journals per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL)

```


```{r}
p.aera.journal.year.pgs.mean = ggplot(d.aera.journal.year.pgs, aes(
  x = publication_date_full, y = page_count_mean,
  label = source_title_shortcode,
  group = source_title_shortcode, color = source_title_shortcode))


p.aera.journal.year.pgs.mean = p.aera.journal.year.pgs.mean + geom_point() + 
  geom_smooth(method = 'loess') + geom_dl(method = 'smart.grid')

p.aera.journal.year.pgs.mean = p.aera.journal.year.pgs.mean + theme_tufte()  + theme(legend.position = 'none')

p.aera.journal.year.pgs.mean + labs(
  title = 'Average published item length is rising',
  subtitle = 'Average number of pages of items published in AERA-journals per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL)

```

# Published items reference counts

```{r}
d.aera.journal.year.ccount = aera %>%
  group_by(publication_date_full, source_title_shortcode) %>%
  summarise(
    cited_reference_count_n = sum(cited_reference_count),
    cited_reference_count_mean = mean(cited_reference_count),
    cited_reference_count_median = median(cited_reference_count),
    cited_references_ppage_mean = mean(references_per_page))

d.aera.journal.year.ccount
```

```{r}
p.aera.journal.year.ccount.mean = ggplot(
  d.aera.journal.year.ccount,
  aes(x = publication_date_full, y = cited_reference_count_mean,
      label = source_title_shortcode,
      group = source_title_shortcode, color = source_title_shortcode))

p.aera.journal.year.ccount.mean = p.aera.journal.year.ccount.mean + geom_point() + 
  geom_smooth(method = 'loess') + geom_dl(method = 'smart.grid')

p.aera.journal.year.ccount.mean = p.aera.journal.year.ccount.mean + theme_tufte()  + theme(legend.position = 'none')

p.aera.journal.year.ccount.mean + labs(
  title = 'The number of references shows different trajectories per AERA-journal',
  subtitle = 'Average number of references in items published in AERA-journals per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL)
```


```{r}
p.aera.journal.year.refppage.mean = ggplot(
  d.aera.journal.year.ccount,
  aes(x = publication_date_full, y = cited_references_ppage_mean,
      label = source_title_shortcode,
      group = source_title_shortcode, color = source_title_shortcode))

p.aera.journal.year.refppage.mean = p.aera.journal.year.refppage.mean + geom_point() + 
  geom_smooth(method = 'loess') + geom_dl(method = 'smart.grid')

p.aera.journal.year.ccount.mean = p.aera.journal.year.refppage.mean + theme_tufte()  + theme(legend.position = 'none')

p.aera.journal.year.ccount.mean + labs(
  title = 'Taking into account increasing item length, shows more modest growth\nin references and the diverging RER-trajectory',
  subtitle = 'Average number of references per page for items published in AERA-journals per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL)
```

# Published items author count

```{r}
d.aera.journal.year.acount = aera %>%
  group_by(publication_date_full, source_title_shortcode) %>%
  summarise(
    author_count_n = sum(author_count),
    author_count_mean = mean(author_count),
    author_count_median = median(author_count))

d.aera.journal.year.acount
```

```{r}
p.aera.journal.year.acount.mean = ggplot(
  d.aera.journal.year.acount,
  aes(x = publication_date_full, y = author_count_mean,
      label = source_title_shortcode,
      group = source_title_shortcode, color = source_title_shortcode))

p.aera.journal.year.acount.mean = p.aera.journal.year.acount.mean + geom_point() + 
  geom_smooth(method = 'loess') + geom_dl(method = 'smart.grid')

p.aera.journal.year.acount.mean = p.aera.journal.year.acount.mean + theme_tufte()  + theme(legend.position = 'none')

p.aera.journal.year.acount.mean + labs(
  title = 'Co- and multi-authored items increase in all AERA-publications',
  subtitle = 'Average number of authors for items published in AERA-journals per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL)
```

