---
title: "AERA journals WOS general descriptives"
output: 
  html_notebook: 
    code_folding: hide
    toc: yes
    toc_float: yes
---

Version: 2018-05-17T21:16:00

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
library(readr)
library(dplyr)
library(here)
library(tidyr)
library(ggplot2)
library(stringr)
library(directlabels)
library(ggthemes)
library(assertr)
library(lubridate)
# library(hrbrthemes)


library(knitr)
opts_chunk$set(echo=FALSE, warning = FALSE, message = FALSE)
```

```{r, warning=FALSE, message=FALSE}
aera_full = read_csv(here::here('data/processed', '20180419_wos_aera_articledata.csv')) %>%
  verify(dim(.) == c(5919, 73))
#spec(aera)
```

# Published item types

In the full downloaded dataset, there are `r nrow(aera_full)` records. 

```{r, message=FALSE, warning=FALSE}
aera_full %>%
  group_by(document_type) %>%
  tally() %>%
  arrange(desc(n)) %>%
  add_tally(n)
```

RF 20/04: Ik denk dat we ons best richten op de volgende 4 document types: Article / Article; Book Chapter / Article; Proceedings Paper / Review. De andere types lijken me niet nodig voor de analyses => Filter 800 items.

```{r}
aera = aera_full %>%
  verify(nrow(.) == 5919) %>%
  filter(document_type %in% c('Article', 'Review', 'Article; Book Chapter', 'Article; Proceedings Paper')) %>%
  verify(nrow(.) == 5119)

aera %>% 
  group_by(document_type) %>%
  tally() %>%
  add_tally(n)
```

Filtered dataset contains `r nrow(aera)` records.

```{r}
aera %>%
  group_by(document_type, source_title_shortcode) %>%
  tally() %>%
  spread(source_title_shortcode, n)
```

```{r}
aera %>%
  group_by(document_type, publication_year) %>%
  tally() %>%
  spread(document_type, n)
```

# Published item count

## Item counts per AERA-journal

```{r}
aera %>%
  group_by(source_title_shortcode) %>%
  tally() %>%
  rename(total_published_items = n)
```

## Item counts per year 

```{r}
aera %>%
  group_by(publication_year) %>%
  tally() %>%
  rename(total_published_items = n)
```

## Item counts per year and AERA-journal

```{r}
aera %>%
  group_by(publication_year, source_title_shortcode) %>%
  tally() %>%
  spread(source_title_shortcode, n)
```

Year-selection: end time series on 2017, i.e. filter records from 2018 (28 records), as incomplete year gives bias.

Filtered dataset contains `r nrow(aera)` records.

```{r}
aera = aera %>%
  verify(nrow(.) == 5119) %>%
  filter(publication_year != 2018) %>% # todo change to character on read-in
  verify(nrow(.) == 5091)
```


```{r, warning=FALSE}
d.aera.journal.year = aera %>%
  group_by(source_title_shortcode, publication_date_full) %>%
  tally()

p.aera.journal.year = ggplot(d.aera.journal.year, aes(
  x = publication_date_full, y = n,
  label = source_title_shortcode,
  group = source_title_shortcode, color = source_title_shortcode))


p.aera.journal.year = p.aera.journal.year + geom_point() + 
  geom_smooth(method = 'loess') + geom_dl(method = 'smart.grid')

p.aera.journal.year = p.aera.journal.year + theme_tufte()  + theme(legend.position = 'none')

p.aera.journal.year + labs(
  title = 'Published item-count is rising, but this is a return to previous levels',
  subtitle = 'Number of items published in AERA-journals per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL
)

# aera %>%
#   group_by(publication_date) %>%
#   tally() %>%
#   mutate()
```

# Published item page count

```{r}
d.aera.journal.year.pgs = aera %>%
  group_by(source_title_shortcode, publication_date_full) %>%
  summarise(
    page_count_n = sum(page_count),
    page_count_mean = mean(page_count),
    page_count_median = median(page_count))

p.aera.journal.year.pgs = ggplot(d.aera.journal.year.pgs, aes(
  x = publication_date_full, y = page_count_n,
  label = source_title_shortcode,
  group = source_title_shortcode, color = source_title_shortcode))


p.aera.journal.year.pgs = p.aera.journal.year.pgs + geom_point() + 
  geom_smooth(method = 'loess') + geom_dl(method = 'smart.grid')

p.aera.journal.year.pgs = p.aera.journal.year.pgs + theme_tufte()  + theme(legend.position = 'none')

p.aera.journal.year.pgs + labs(
  title = 'Total number of pages published is generally risng',
  subtitle = 'Number of pages published in AERA-journals per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL)

```


```{r}
p.aera.journal.year.pgs.mean = ggplot(d.aera.journal.year.pgs, aes(
  x = publication_date_full, y = page_count_mean,
  label = source_title_shortcode,
  group = source_title_shortcode, color = source_title_shortcode))


p.aera.journal.year.pgs.mean = p.aera.journal.year.pgs.mean + geom_point() + 
  geom_smooth(method = 'loess') + geom_dl(method = 'smart.grid')

p.aera.journal.year.pgs.mean = p.aera.journal.year.pgs.mean + theme_tufte()  + theme(legend.position = 'none')

p.aera.journal.year.pgs.mean + labs(
  title = 'Average published item length is rising',
  subtitle = 'Average number of pages of items published in AERA-journals per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL)

```

# Published items reference counts

```{r}
d.aera.journal.year.ccount = aera %>%
  group_by(publication_date_full, source_title_shortcode) %>%
  summarise(
    cited_reference_count_n = sum(cited_reference_count),
    cited_reference_count_mean = mean(cited_reference_count),
    cited_reference_count_median = median(cited_reference_count),
    cited_references_ppage_mean = mean(references_per_page))

d.aera.journal.year.ccount
```

## Trend for all AERA-journals

```{r}
p.aera.journal.year.ccount.mean = ggplot(
  d.aera.journal.year.ccount,
  aes(x = publication_date_full, y = cited_reference_count_mean,
      label = source_title_shortcode,
      group = source_title_shortcode, color = source_title_shortcode))

p.aera.journal.year.ccount.mean = p.aera.journal.year.ccount.mean + geom_point() + 
  geom_smooth(method = 'loess') + geom_dl(method = 'smart.grid')

p.aera.journal.year.ccount.mean = p.aera.journal.year.ccount.mean + theme_tufte()  + theme(legend.position = 'none')

p.aera.journal.year.ccount.mean = p.aera.journal.year.ccount.mean + labs(
  title = 'The number of references shows different trajectories per AERA-journal',
  subtitle = 'Average number of references in items published in AERA-journals per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL)

p.aera.journal.year.ccount.mean
```


```{r}
p.aera.journal.year.refppage.mean = ggplot(
  d.aera.journal.year.ccount,
  aes(x = publication_date_full, y = cited_references_ppage_mean,
      label = source_title_shortcode,
      group = source_title_shortcode, color = source_title_shortcode))

p.aera.journal.year.refppage.mean = p.aera.journal.year.refppage.mean + geom_point() + 
  geom_smooth(method = 'loess') + geom_dl(method = 'smart.grid')

p.aera.journal.year.ccount.mean = p.aera.journal.year.refppage.mean + theme_tufte()  + theme(legend.position = 'none')

p.aera.journal.year.ccount.mean + labs(
  title = 'Taking into account increasing item length, shows more modest growth\nin references and the diverging RER-trajectory',
  subtitle = 'Average number of references per page for items published in AERA-journals per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL)
```

## Trend for AERJ and RER

RF, 17/05: De “published items reference counts” grafiek bevat nu de data van alle AERA tijdschriften. Maar kan je die gegevens voor RER en AERJ apart uitzetten in twee grafieken, naar analogie met een oude grafiek over het gemiddeld aantal auteurs (zie hiervoor bijlage)?

```{r}
d.aera.journal.year.ccount.rer_aerj = d.aera.journal.year.ccount %>%
  filter(source_title_shortcode %in% c('RER', 'AERJ'))
```


```{r}
p.aera.journal.year.ccount.mean.rer_aerj = ggplot(
  d.aera.journal.year.ccount.rer_aerj,
  aes(x = publication_date_full, y = cited_reference_count_mean,
      label = source_title_shortcode,
      group = source_title_shortcode, color = source_title_shortcode))

p.aera.journal.year.ccount.mean.rer_aerj = p.aera.journal.year.ccount.mean.rer_aerj +
  geom_point() + 
  geom_smooth(method = 'loess') + facet_grid(source_title_shortcode ~ .)

p.aera.journal.year.ccount.mean.rer_aerj = p.aera.journal.year.ccount.mean.rer_aerj +
  # theme_tufte() + 
  theme(
    panel.grid.minor = element_blank(),
    legend.position = 'none')

p.aera.journal.year.ccount.mean.rer_aerj = p.aera.journal.year.ccount.mean.rer_aerj + labs(
  title = 'Both AERJ and RER show an increase in the average number of references ',
  subtitle = 'Average number of references in items published per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL)

p.aera.journal.year.ccount.mean.rer_aerj = p.aera.journal.year.ccount.mean.rer_aerj + 
  scale_y_continuous(breaks = c(10, 50, 100, 150)) +
  scale_x_date(breaks = c(ymd('1956-06-15'), ymd('1964-06-15'),
                          ymd('1970-06-15'), ymd('1980-06-15'), 
                          ymd('1990-06-15'), ymd('2000-06-15'), 
                          ymd('2010-06-15'), ymd('2017-06-15')),
               labels = c('1956', '1964', '1970', '1980', '1990', '2000', '2010', '2017'))

p.aera.journal.year.ccount.mean.rer_aerj
```

```{r}
d.aera.journal.year.ccount.rer_aerj %>%
  # filter(source_title_shortcode == 'AERJ') %>%
  group_by(source_title_shortcode) %>%
  summarise(
    min_year = min(publication_date_full),
    min_count_mean = min(cited_reference_count_mean))
# min(d.aera.journal.year.ccount.rer_aerj$publication_date_full)
# max(d.aera.journal.year.ccount.rer_aerj$publication_date_full)
```


```{r}
p.aera.journal.year.refppage.mean.rer_aerj = ggplot(
  d.aera.journal.year.ccount.rer_aerj,
  aes(x = publication_date_full, y = cited_references_ppage_mean,
      label = source_title_shortcode,
      group = source_title_shortcode, color = source_title_shortcode))

p.aera.journal.year.refppage.mean.rer_aerj = p.aera.journal.year.refppage.mean.rer_aerj +
  geom_point() + 
  geom_smooth(method = 'loess') + facet_grid(source_title_shortcode ~ .)

p.aera.journal.year.refppage.mean.rer_aerj = p.aera.journal.year.refppage.mean.rer_aerj +
  # theme_tufte()  +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = 'none')

p.aera.journal.year.refppage.mean.rer_aerj = p.aera.journal.year.refppage.mean.rer_aerj + 
  scale_x_date(breaks = c(ymd('1956-06-15'), ymd('1964-06-15'),
                          ymd('1970-06-15'), ymd('1980-06-15'), 
                          ymd('1990-06-15'), ymd('2000-06-15'), 
                          ymd('2010-06-15'), ymd('2017-06-15')),
               labels = c('1956', '1964', '1970', '1980', '1990', '2000', '2010', '2017'))

p.aera.journal.year.refppage.mean.rer_aerj = p.aera.journal.year.refppage.mean.rer_aerj +
  labs(
    title = 'Taking into account increasing item length, shows more modest growth\nin references and the diverging RER-trajectory',
    subtitle = 'Average number of references per page for items published per year, 1956-2017',
    caption = 'Web of Knowledge, queried 2018-04-19',
    x = NULL, y = NULL)
p.aera.journal.year.refppage.mean.rer_aerj
```

```{r}
# save version with and without titles
ggsave(
  here::here('describe/figures', 'aera_wos_refcount_mean_rer_aerj_1956-2017_title.png'),
  p.aera.journal.year.ccount.mean.rer_aerj, width = 8, height = 6)
ggsave(
  here::here('describe/figures', 'aera_wos_refcount_mean_rer_aerj_1956-2017_notitle.png'),
  p.aera.journal.year.ccount.mean.rer_aerj + labs(
    title = NULL, subtitle = NULL, caption = NULL),
  width = 8, height = 6)

ggsave(
  here::here('describe/figures', 'aera_wos_refcount_mean_pp_rer_aerj_1956-2017_title.png'),
  p.aera.journal.year.refppage.mean.rer_aerj, width = 8, height = 6)
ggsave(
  here::here('describe/figures', 'aera_wos_refcount_mean_pp_rer_aerj_1956-2017_notitle.png'),
  p.aera.journal.year.refppage.mean.rer_aerj + labs(
    title = NULL, subtitle = NULL, caption = NULL),
  width = 8, height = 6)
```



# Published items author count

```{r}
d.aera.journal.year.acount = aera %>%
  group_by(publication_date_full, source_title_shortcode) %>%
  summarise(
    author_count_n = sum(author_count),
    author_count_mean = mean(author_count),
    author_count_median = median(author_count))

d.aera.journal.year.acount
```

```{r}
p.aera.journal.year.acount.mean = ggplot(
  d.aera.journal.year.acount,
  aes(x = publication_date_full, y = author_count_mean,
      label = source_title_shortcode,
      group = source_title_shortcode, color = source_title_shortcode))

p.aera.journal.year.acount.mean = p.aera.journal.year.acount.mean + geom_point() + 
  geom_smooth(method = 'loess') + geom_dl(method = 'smart.grid')

p.aera.journal.year.acount.mean = p.aera.journal.year.acount.mean + theme_tufte()  + theme(legend.position = 'none')

p.aera.journal.year.acount.mean + labs(
  title = 'Co- and multi-authored items increase in all AERA-publications',
  subtitle = 'Average number of authors for items published in AERA-journals per year, 1956-2017',
  caption = 'Web of Knowledge, queried 2018-04-19',
  x = NULL, y = NULL)
```

